The role of advection versus diffusion in intraventricular transport

### 2025-03-10 paper/main.tex
Results 5 and 6

\begin{figure}%[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{
    \textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$ 
    in each region of interest (ROI)
    observed during experiments of photoconversion of Dendra2 proteins in ROI 1.
    \textbf{b.} Mean concentration $\cbar(t)$ in each ROI in a
    simulation of Dendra2 protein transport with
    the diffusion coefficient $D_3$.
    \textbf{c.} Time to $\Delta F(t)$ reaches a threshold value of 0.25 in
    one representative zebrafish larva.
    Time to $\cbar(t)$ reaches a threshold value of 0.25 for
    a transport simulation with diffusion coefficient $D_3$.
    We have excluded the posterior-most regions of the geometries,
    because the values of $\Delta F$ and $\cbar$ never exceeded the threshold
    value of 0.25 in these regions.}
    \label{fig:fig3}
\end{figure}
% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6
% Final mean c values ROIs 2-6: 0.7637806  0.72025859 0.81903216 0.56204164 0.34926411

@@

We next mimic the \emph{smh} mutants in the computational model by
considering transport of Dendra2 with CSF flow driven by cardiac
pressure pulsations only, representing the absence of motile
cilia. Comparing the in-silico mutants with transport predictions in
the presence of both cardiac pulsations and motile cilia, we observe
noticeably delayed dynamics in ROIs 2, 4 and 5, but little change for
ROIs 3 and 6 (\Crefpanels{fig:fig4}{c}).

\mer{MER: MER and HH stopped here.}

Comparing with the experiment, the difference between the control and
mutant scenarios was even greater for the simulations, with the
times-to-threshold in ROI 2 and 4 being 90\% and 34\% higher in the
absence of cilia-mediated flow. Moreover, ROI 5 exhibits very
different $\cbar$ profiles for the two configurations in our
computational model, but not in our experimental data.  Altogether,
our findings identify that diffusion plays an important role for
solute movement to regions upstream in the cilia flow fields or to
distant ventricles.

%% Again, the simulated Dendra2 distribution patterns differ from the

%% Next, to identify whether our model would recapitulate these findings,
%% we simulate CSF flow driven only by cardiac pressure pulsations and
%% the resulting diffusion and advection of

%% Dendra2 transport

%% by using the only-cardiac/no-cilia flow model 
%% and compare it with the baseline model
%% (both cilia and cardiac motion) (\Crefpanels{fig:fig4}{c}).

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F(t)$
    in each region of interest (ROI) for control (black lines)
    and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets. The $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%, with 
    the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}

\subsection{Loss of cilia motility in specific subpopulation slows down transport}
We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the role of these lineages on CSF transport, we simulate transport
upon region-specific cilia paralysis by modifying the boundary conditions
of the CSF flow model. Three cilia-covered regions were considered separately:
the telencephalic (anterior) ventricle, and the dorsal and ventral parts
of the diencephalic (middle) ventricle (\Crefpanels{fig:fig5}{a}).
The cilia tangential traction boundary condition is for each case
changed to a free-slip boundary condition in the paralyzed region,
prescribing zero tangential forces to
represent absence of cilia motion.
We simulate CSF flow modeled by the Stokes equations with the
modified boundary conditions and use the resulting
velocity fields in transport simulations.

@@

Changes in the transport dynamics of Dendra2 are most prominent when removing the dorsal cilia
in the diencephalic ventricle. All ROIs (except ROI 1, where we impose the photoconversion curve)
exhibit changes in the dynamics of $\cbar$ (\Crefpanels{fig:fig5}{c}).
Compared to the baseline model results, we observe a slower increase and lower final concentration
in the middle and anterior ventricles (ROIs 2--5),
and more rapid transport to the posterior ventricle (ROI 6).
In contrast, removal of telencephalic cilia does not alter the evolution of the
mean concentration $\cbar$ with time (\Crefpanels{fig:fig5}{c}).
Finally, without ventral cilia in the diencephalic ventricle,
more solute spreads towards the anterior ventricle (ROI 5) and
less solute spreads towards the posterior ventricle (ROI 6).
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate ciliated regions cilia paralysis is simulated
    by imposing a zero tangential traction boundary condition in the respective region.
    \textbf{b.} The total tangential force applied with the tangential traction
    boundary condition in the three separate ciliated regions, as well as the total
    (the sum of the three regions), calculated with \Cref{eq:tangential_force_calculation}. 
    \textbf{c.} Mean concentration $\cbar(t)$ for the six regions of interest (ROIs) in the three 
    paralysis scenarios considered. The region named in the legend is the region
    where we modify the cilia tangential traction boundary condition to a free-slip condition
    in the CSF flow model.}
    \label{fig:fig5}
\end{figure}

\subsection{Modification of ventricles geometry impacts protein distribution}
Variations in ventricular morphology impact CSF flow patterns.
To assess how ventricular morphologies influence solute distribution,
we consider four alternative brain ventricles meshes:
one where the duct connecting the anterior and middle ventricles is constricted
(66 \% reduction in cross-sectional area);
one where the duct connecting the middle and posterior ventricles is constricted
(33 \% reduction);
one where the middle ventricle is shrunk in the lateral direction (43 \% reduction);
and the combination of the above three modifications (\Crefpanels{fig:fig6}{a}). In these
four modified geometries the volume is reduced by 3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively,
compared to the original geometry.
We simulate CSF flow in the modified geometries with the baseline flow model
and use the resulting velocity field to simulate transport of Dendra2
(\Crefpanels{fig:fig6}{b}).

Constricting the connection between the anterior and middle ventricle leads to
slower transport to the anterior ventricle (ROI 5),
and a slightly higher mean concentration in ROIs 2--4 and 6, compared to the original geometry
(\Crefpanels{fig:fig6}{c}, shrunk fore-mid scenario). This is notwithstanding a
9.5\% reduction in ventral cilia force magnitude
(\Crefpanels{fig:fig6}{d}), which are the cilia that propel CSF and solute towards ROI 6.
On the other hand, constricting the duct connecting the middle and the posterior ventricle
leads to less transport into the posterior ventricle (ROI 6),
and slightly more transport towards the anterior ventricle (ROI 5)
(\Crefpanels{fig:fig6}{c}, shrunk mid-hind scenario).
Faster transport to the anterior ventricle (ROI 5) and slower transport to the
posterior ventricle (ROI 6) is also the case for the geometry with a
shrunk middle ventricle.
Here also the change in transport to ROI 6 is counter to expectation when comparing with
ventral cilia forces, with transport reducing in spite of a 6.6\% increase in the
ventral cilia forces.
Additionally, for the shrunk middle configuration, the transport upstream of CSF flow towards the
ventral-posterior region of the middle ventricle (ROI 3) is slowed down.
For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have lower mean concentrations.
Altogether, our results show that the geometry and width of the ventricular ducts matter
for solute transport, with notable differences in transport from relatively small
geometry modifications. This may explain why we observed larger differences
in the cilia-deficient computational model compared to the experimental data in
\Cref{fig:fig4}, as the model and the fish have different duct diameters.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios thus
    display $F_i/F_{\mathrm{original}}\cdot 100\%$.}
    \label{fig:fig6}
\end{figure}

@@

Meanwhile, we identified that transport of larger molecules
(thus with smaller diffusion coefficients),
such as 150 nm exosomes, are almost entirely dependent on advection.
Indeed, there was essentially zero solute in the posterior ventricle at the
final simulation time (around 850 s), a ventricle where transport is largely driven
by diffusion due to the absence of motile cilia.
For $D_1$, an estimate of the diffusion timescale $t_d=L^2/D_1$ over a distance
$L=200$ \textmu m from the middle to the posterior ventricle is approximately 24500 s.
Advective transport by the mean velocity $\overline{U}=1.41 \ \mu$m/s over the same
distance would happen on a timescale of the order of $t_a=L/\overline{U}=141$ s.
Comparing this with the minuscule values of $\cbar_6$ after simulating transport
with $D_1$ for 850 s, one can conclude that if advection
played a role in transporting solute from the photoconversion site in the middle ventricle
to the posterior ventricle, values of $\cbar_6$ would be higher than those observed.

@@

\subsection{Computational model assumptions and boundary conditions}
To model CSF, we employed slip boundary conditions at the ventricular walls,
contrasting the classical choice of a noslip boundary condition.
The use of slip boundary conditions is motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles, both previously and in this work, in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. Slip boundary
conditions allow for this behavior, in contrast to a noslip condition where the velocity
has to go to zero at the wall.

As opposed to a couple of previous continuum model simulation studies of
cilia-mediated flow where cilia were represented by body
forces,~\cite{Siyahhan2014FlowVentricles, Thouvenin2020OriginCanal}
we chose to model the cilia as a force
exerted tangentially on the boundary.
We deem this a good representation of the real nature of cilia,
since they are microscale structures confined to the ventricular walls that
generate flow through collective motion. Furthermore, it avoids the need of mesh
refinement near the ciliated boundaries in the case that a
body force were to be applied in a cilia layer of depth around
5 \textmu m~\cite{Salman2022ComputationalEmbryo}.

@@

To simulate protein transport in the brain ventricles, we numerically solved
an advection-diffusion equation. Except for where we imposed a normal pressure
boundary condition on the CSF flow,
we imposed a no-flux condition on the concentration. Imposing a pressure-driven
cardiac-related flow on certain parts of the CSF flow boundary leads to in- and outflow,
and the advective transport then requires handling of in- and outflux boundary conditions.
We set boundary conditions such that the total mass is conserved in the
limit that the timestep size is zero, reflecting our assumption that the
ventricular system is closed.
We handled the outflux by leaving the advective flux term as an unknown in the weak formulation
of the advection-diffusion equation. This has been referred to with different names
in previous works, names such as the ``free boundary condition"~\cite{Papanastasiou1992ACondition},
the ``no boundary condition"~\cite{Griffiths1997TheCondition}, and the 
``consistent flux boundary condition"~\cite{Lynch2020NumericalHemodynamics}. 
In leaving the advective flux term in the weak form we are lacking a boundary condition,
which is problematic in the continuous sense of the problem.
However, the discrete problem is well-posed, because the approach represents
constraining higher-order derivatives of the concentration~\cite{Griffiths1997TheCondition}.

Our approach can be interpreted as an approximate periodic boundary condition, where we set
the inflow boundary flux at timestep $n$ equal to the outflow boundary flux at the previous
timestep $n-1$. This is expected to be of low-order accuracy, and for high Péclet numbers,
the boundary condition affects the concentration field in the vicinity of the
in- and outflow boundaries
due to large advective fluxes. For the simulation scenarios we considered,
both in terms of model parameters
and time span, the photoconversion region where solute originates
is far enough away from the in- and outflow
boundaries for the flux boundary condition to not significantly impact the results. Nevertheless,
we deliberately excluded the space proximal to in- and outflow boundaries in
the anterior and posterior regions of interest (5 and 6) to limit adverse effects.

@@

\subsection{Numerical approximation and convergence rates}
The Brezzi-Douglas-Marini and discontinuous Galerkin discretization scheme
($\mathrm{BDM}_1$-$\mathrm{DG}_0$) that we employ for the Stokes equations
is computationally more expensive than a more conventional approach based on
e.g. Taylor-Hood discretization~\cite{Stenberg1990ErrorProblem}
($P_2-P_1$ elements) or lower-order enriched
elements~\cite{Brezzi2011MixedMethods},
such as the mini-element ($P_1-P_1$ with a velocity stabilization).
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property,
since numerical experiments proved that a divergence-free velocity field was
fundamental to attain a stable discretization of the advection-diffusion equation
in advection-dominated regions. Using conventional
Taylor-Hood elements for the Stokes equations
stability of the advection-diffusion discretization either
required excessive mesh refinement and/or problem-specific tuning of
hyperparameters when employing a Streamline-Upwind/Petrov Galerkin stabilization
method~\cite{Franca1992StabilizedModel}, which is why we also opted for a DG method
for the advection-diffusion equation.
Probing numerical values of the divergence of
the velocity when using Taylor-Hood elements for the fluid equations revealed that,
on the standard computational mesh, the $L^2$ norm of $\nabla\cdot\uu_h$ was
in the order of $10^{-6}$, while the
maximum value of the divergence of the velocity was of order unity. This is a result
of the fact that incompressibility is only achieved in a (global) weak sense when
employing Taylor-Hood elements. On a refined mesh with roughly six times as many
computational cells as the standard mesh, the numerical values of
$\normltwo{\nabla\cdot\uu_h}$ and $\normlinf{\nabla\cdot\uu_h}$ were still
of order unity. This may lead to the
hyperbolic advective term in the advection-diffusion equation accumulating errors
that grow unstable over time.

A priori error estimates for the $\mathrm{BDM}_k$-$\mathrm{DG}_{k-1}$ scheme are naturally 
derived in a mesh-dependent broken $H^k$-norm for $\uu$ and the $L^2$-norm for the pressure.
The respective error norms are expected to decay with order $k$~\cite{Cockburn2005AEquations}.
For the sake of simplicity,~\Cref{tab:cylinder_error_rates}
reports the velocity errors in the standard $H^k$ norm. Using polynomial degree  
$k=1$, we observe linear convergence in both the velocity and the pressure.   
\mk{MK: I see 1.5 in the table for $\uu$ in $L^2$?}
\lyng{HH: Yes, we should expect 2 there? Maybe be more specific in the comment:
"Using polynomial degree $k=1$, we observe linear convergence in both the velocity $H^1$ error
and the pressure $L^2$ error. For the $L^2$ error of the velocity, we observe a suboptimal
convergence order of 1.5"}

@@

%\section{Conclusion}
In conclusion, we have presented a numerical model
of cerebrospinal fluid flow and solute transport in
zebrafish brain ventricles based on finite element methods.
A realistic ventricular geometry from medical imaging was used to perform a
computational study. The flow model reproduces
the experimentally observed flow features of cilia-mediated cerebrospinal fluid
flow in zebrafish brain ventricles. Coupling the flow to the transport model,
we estimated how slowly or rapidly molecules of varying size are transported in
the ventricles. 
The numerical model was validated with in vivo experiments, measuring the movement of
a photoconverted fluorescent protein in zebrafish brain ventricles,
both for healthy animals and mutant animals with ciliary dysfunction.
We observed good alignment between computational and experimental results.
Simulations indicate that cilia-driven advection is important
for transport of larger molecules, whereas diffusion accounts for most of
the transport of smaller molecules. 
Experimental results support the conclusion that cilia only affect the
local distribution of small proteins,
while most of their transport is governed by diffusion.

@@

To evaluate the relative importance of advection and
diffusion as transport mechanisms, we consider the Péclet number
\begin{equation*}
    \mathrm{Pe} = \frac{\mathrm{diffusion \ timescale}}{\mathrm{advection \ timescale}}
    = \frac{t_d}{t_a} = \frac{L_p^2/D}{L_p/U_p} = \frac{L_p U_p}{D}.
\end{equation*}
Thus, if $\mathrm{Pe} > 1$, transport is dominated by advection,
and if $\mathrm{Pe} < 1$ transport is dominated by diffusion.
Here, $L_p$ and $U_p$ are characteristic length and velocity scales,
respectively, and $D$ is the solute diffusion coefficient.
In calculating a global Péclet number, we use $L_p=600$ \textmu m, the approximate length of the
ventricular geometry along the rostrocaudal ($x$-) axis.
We use a mean velocity $\overline{U}$ as the characteristic velocity $U_p$.
The mean velocity $\overline{U}$ is calculated by first averaging the velocity in space:
\begin{equation*}
    \overline{u}^2 = \frac{1}{|\Omega|}\int_{\Omega}\uu\cdot\uu\dx,
\end{equation*}
and then averaging $\overline{u}$ in time over one cardiac cycle to yield $\overline{U} = \fixme{X}$ \fixme{unit}.
To assess the assumption that inertial forces are negligible compared to
viscous forces, we also compute the Reynolds number
\begin{equation*}
    \mathrm{Re} = \frac{\mathrm{inertial \ forces}}{\mathrm{viscous \ forces}} = \frac{\rho U_r L_r}{\mu},
\end{equation*}
with $L_r = \fixme{X}$ \fixme{unit} and $U_r = \fixme{X}$ \fixme{unit}
as length and velocity scales characteristic of the flow.

>>>>

\begin{figure}%[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{
    \textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$ 
    in each region of interest (ROI)
    observed during experiments of photoconversion of Dendra2 proteins in ROI 1. 
    Bold lines are mean values of the cohort and shaded regions cover one standard devation errors
    in the mean.
    \textbf{b.} Mean concentration $\cbar(t)$ in each ROI in a
    simulation of Dendra2 protein transport with
    the diffusion coefficient $D_3$.
    \textbf{c.} Time to $\Delta F(t)$ reaches a threshold value of 0.25 in
    one representative zebrafish larva.
    Time to $\cbar(t)$ reaches a threshold value of 0.25 for
    a transport simulation with diffusion coefficient $D_3$.
    We have excluded the posterior-most regions of the geometries,
    because the values of $\Delta F$ and $\cbar$ never exceeded the threshold
    value of 0.25 in these regions.}
    \label{fig:fig3}
\end{figure}
% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6
% Final mean c values ROIs 2-6: 0.7637806  0.72025859 0.81903216 0.56204164 0.34926411

@@

The \emph{smh} mutants are mimicked in the computational model by
considering transport of Dendra2 with CSF flow driven by cardiac
pressure pulsations only, representing the absence of motile
cilia. Comparing the in-silico mutants with transport predictions in
the presence of both cardiac pulsations and motile cilia, we observe
noticeably delayed dynamics in ROIs 2, 4 and 5, but little change for
ROIs 3 and 6 (\Crefpanels{fig:fig4}{c}).

Compared to experimental data, the difference between the control and
mutant scenarios was even greater for the simulations, with the
times-to-threshold in ROI 2 and 4 being 90\% and 34\% higher in the
absence of cilia-mediated flow. Moreover, the ROI 5 $\cbar$ profiles
for the two configurations in our computational model deviate,
which is not the case in the experiments.  Altogether,
our findings identify that, for smaller proteins such as Dendra2,
diffusion plays an important role for
solute movement to regions upstream in the cilia flow fields or to
distant ventricles, while cilia affect the downstream mean concentrations.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F(t)$
    in each region of interest (ROI) for control (black lines, $n=8$)
    and \emph{smh} mutant (red lines, $n=6$) fish. 
    Bold lines are mean values of the cohorts and shaded regions cover one standard devation errors
    in the means.
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets. The $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%, with 
    the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}

\subsection{Loss of cilia motility in specific subpopulation slows down transport}

We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the effects of these lineages on CSF flow and transport, 
we investigate region-specific cilia paralysis with the computational model.
Three cilia-covered regions were considered paralyzed separately:
the telencephalic (anterior) ventricle, and the dorsal and ventral parts
of the diencephalic (middle) ventricle (\Crefpanels{fig:fig5}{a}).
To simulate cilia paralysis, the cilia tangential traction boundary condition
is for each case changed to a free-slip boundary condition
in the paralyzed region,
prescribing zero tangential forces to
represent absence of cilia motion.

@@

Changes in the transport of Dendra2 are most prominent when removing the dorsal cilia
in the diencephalic ventricle. The mean concentration profiles 
of all ROIs (except ROI 1, where we impose the photoconversion curve)
exhibit changes (\Crefpanels{fig:fig5}{c}).
Compared to the baseline model results, we observe less solute distribution
in the middle and anterior ventricles (ROIs 2--5),
and more rapid transport to the posterior ventricle (ROI 6).
Meanwhile, removal of telencephalic cilia does not alter the 
mean concentrations (\Crefpanels{fig:fig5}{c}).
Finally, without ventral cilia in the diencephalic ventricle,
more solute spreads towards the anterior ventricle (ROI 5) and
less solute spreads towards the posterior ventricle (ROI 6).
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate ciliated regions cilia paralysis is simulated
    by imposing a zero tangential traction boundary condition in the respective region.
    \textbf{b.} The total tangential force applied with the tangential traction
    boundary condition in the three separate ciliated regions, as well as the total
    (the sum of the three regions), calculated with \Cref{eq:tangential_force_calculation}. 
    \textbf{c.} Mean concentration $\cbar(t)$ for the six regions of interest (ROIs) in the three 
    paralysis scenarios considered. The region named in the legend is the region
    where we modify the cilia tangential traction boundary condition to a free-slip condition
    in the CSF flow model.}
    \label{fig:fig5}
\end{figure}

\subsection{Modification of ventricles geometry impacts protein distribution}
Variations in ventricular morphology impact CSF flow patterns.
To assess how these variations influence solute distribution,
we consider four alternative brain ventricles meshes:
one where the duct connecting the anterior and middle ventricles is constricted
(66 \% reduction in cross-sectional area);
one where the duct connecting the middle and posterior ventricles is constricted
(33 \% reduction);
one where the middle ventricle is shrunk in the lateral direction (43 \% reduction);
and the combination of the above three modifications (\Crefpanels{fig:fig6}{a, b}). In these
four modified geometries the volume is reduced by 3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively,
compared to the original geometry.

Constricting the connection between the anterior and middle ventricle leads to
slower transport to the anterior ventricle (ROI 5),
and a slightly higher mean concentration in ROIs 2--4 and 6, compared to the original geometry
(\Crefpanels{fig:fig6}{c}, shrunk fore-mid scenario). This is notwithstanding a
9.5\% reduction in ventral cilia force magnitude
(\Crefpanels{fig:fig6}{d}), which are the cilia that propel CSF and solute towards ROI 6.
Constricting the mid-hind brain connection
leads to less transport into the posterior ventricle (ROI 6),
and slightly more transport towards the anterior ventricle (ROI 5).
Faster transport to the anterior ventricle (ROI 5) and slower transport to the
posterior ventricle (ROI 6) is also the case for the geometry with a
shrunk middle ventricle.
Here the transport to the posterior ventricle (ROI 6) is reduced,
in spite of a 6.6\% increase in the ventral cilia forces.
Additionally, for the shrunk middle configuration, the transport upstream of CSF flow towards the
ventral-posterior region of the middle ventricle (ROI 3) is slowed down.

For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have reduced
mean concentrations. Altogether, our results show that the geometry and
width of the ventricular ducts matter
for solute transport, with notable differences in transport from relatively small
geometry modifications. This may explain why we observed larger differences
in the cilia-deficient computational model compared to the experimental data in
\Cref{fig:fig4}, as the model and the fish have different duct diameters.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios thus
    display $F_i/F_{\mathrm{original}}\cdot 100\%$.}
    \label{fig:fig6}
\end{figure}

@@

Meanwhile, we identified that transport of larger molecules
(thus with smaller diffusion coefficients),
such as 150 nm exosomes, are almost entirely dependent on advection.
Indeed, there was essentially zero solute in the posterior ventricle at the
final simulation time (around 860 s), a ventricle where transport is largely driven
by diffusion due to the absence of motile cilia.
For $D_1$, an estimate of the diffusion timescale $t_d=L^2/D_1$ over a distance
$L=200$ \textmu m from the middle to the posterior ventricle is approximately 24500 s.
Advective transport by the mean velocity $\overline{U}=2.4 \ \mu$m/s over the same
distance would happen on a timescale of the order of $t_a=L/\overline{U}=83$ s.
Comparing this with the minuscule values of $\cbar_6$ after simulating transport
with $D_1$ for 850 s, one can conclude that if advection
played a role in transporting solute from the photoconversion site in the middle ventricle
to the posterior ventricle, values of $\cbar_6$ would be higher than those observed.

@@

\subsection{Computational modeling aspects}
We chose to model the cilia as a force
exerted tangentially on the boundary, as opposed to a
couple of previous continuum model simulation studies of
cilia-mediated flow where cilia were represented by body
forces,~\cite{Siyahhan2014FlowVentricles, Thouvenin2020OriginCanal}.
We employed slip boundary conditions at the ventricular walls,
motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles, both previously and in this work, in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
In addition to its favorable mass-conserving property, the 
Brezzi-Douglas-Marini and discontinuous Galerkin (BDM-DG) discretization scheme 
we employed for the Stokes equations
lets us impose the impermeability condition strongly, and the tangential
traction boundary condition weakly, without any tailored approach,
since these boundary conditions appear naturally in the weak form of the
Stokes equations~(\Cref{subsec:appendixA1}). If instead e.g. Taylor-Hood
elements were to be used, the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}.

The BDM-DG scheme ($\mathrm{BDM}_1$-$\mathrm{DG}_0$) that we employed for the Stokes equations
is computationally more expensive than a more conventional approach based on
e.g. Taylor-Hood discretization~\cite{Stenberg1990ErrorProblem}
($P_2-P_1$ elements) or lower-order enriched
elements~\cite{Brezzi2011MixedMethods},
such as the mini-element ($P_1-P_1$ with a velocity stabilization).
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property,
as the elements by definition satisfy the incompressibility constraint for the fluid.
This leads to $\normlinf{\nabla\cdot\uu_h}$ being equal to zero to within solver tolerances,
an important property for the stability of the advection-diffusion equation\cite{CITATION}.

@@

To simulate protein transport in the brain ventricles, we numerically solved
an advection-diffusion equation. Except for where we imposed a normal pressure
boundary condition on the CSF flow,
we imposed a no-flux condition on the concentration. Imposing a pressure-driven
cardiac-related flow on certain parts of the CSF flow boundary leads to in- and outflow,
and the advective transport then requires handling of in- and outflux boundary conditions.
We set an approximate periodic boundary condition on the the in- and outflux boundaries,
such that the total mass is conserved in the
limit that the timestep size is zero, reflecting our assumption that the
ventricular system is closed.
We handled the outflux by leaving the advective flux term as an unknown in the weak formulation
of the advection-diffusion equation. This has been referred to with different names
in previous works, names such as the ``free boundary condition"~\cite{Papanastasiou1992ACondition},
the ``no boundary condition"~\cite{Griffiths1997TheCondition}, and the 
``consistent flux boundary condition"~\cite{Lynch2020NumericalHemodynamics}. 
In leaving the advective flux term in the weak form we are lacking a boundary condition,
which is problematic in the continuous sense of the problem.
However, the discrete problem is well-posed, because the approach represents
constraining higher-order derivatives of the concentration~\cite{Griffiths1997TheCondition}. 
We deliberately excluded the space proximal to in- and outflow boundaries in
the anterior and posterior regions of interest (5 and 6) to limit adverse effects of 
the advective transport at the in- and outflux boundaries on the mean concentrations.

@@

@@

In conclusion, we have presented a numerical model
of cerebrospinal fluid flow and solute transport in
zebrafish brain ventricles based on finite element methods.
A realistic ventricular geometry from medical imaging was used to perform a
computational study. The flow model reproduces
the experimentally observed flow features of cilia-mediated cerebrospinal fluid
flow in zebrafish brain ventricles. Coupling the flow to the transport model,
we estimated how slowly or rapidly molecules of varying size are transported in
the ventricles. 
The numerical model was validated with in vivo experiments, measuring the movement of
a photoconverted fluorescent protein in zebrafish brain ventricles,
both for healthy animals and mutant animals with ciliary dysfunction.
We observed good alignment between computational and experimental results.
Simulations indicate that cilia-driven advection is important
for transport of larger molecules, whereas diffusion accounts for most of
the transport of smaller molecules. 
Experimental results support the conclusion that cilia only affect the
local distribution of small proteins,
while most of their transport is governed by diffusion.

@@

To evaluate the relative importance of advection and
diffusion as transport mechanisms, we consider the Péclet number
\begin{equation*}
    \mathrm{Pe} = \frac{\mathrm{diffusion \ timescale}}{\mathrm{advection \ timescale}}
    = \frac{t_d}{t_a} = \frac{L_p^2/D}{L_p/U_p} = \frac{L_p U_p}{D}.
\end{equation*}
Thus, if $\mathrm{Pe} > 1$, transport is dominated by advection,
and if $\mathrm{Pe} < 1$ transport is dominated by diffusion.
Here, $L_p$ and $U_p$ are characteristic length and velocity scales,
respectively, and $D$ is the solute diffusion coefficient.
In calculating a global Péclet number, we use $L_p=600$ \textmu m, the approximate length of the
ventricular geometry along the rostrocaudal ($x$-) axis.
We use a mean velocity $\overline{U}$ as the characteristic velocity $U_p$.
The mean velocity $\overline{U}$ is calculated by first averaging the velocity in space:
\begin{equation*}
    \overline{u}^2 = \frac{1}{|\Omega|}\int_{\Omega}\uu\cdot\uu\dx,
\end{equation*}
and then averaging $\overline{u}$ in time over one cardiac cycle to yield $\overline{U} = 2.4$ \textmu m/s.
To assess the assumption that inertial forces are negligible compared to
viscous forces, we also compute the Reynolds number
\begin{equation*}
    \mathrm{Re} = \frac{\mathrm{inertial \ forces}}{\mathrm{viscous \ forces}} = \frac{\rho U_r L_r}{\mu},
\end{equation*}
with $L_r = 110$ \textmu m and $U_r = 27.7$ \textmu m/s
as length and velocity scales characteristic of the flow.
